%%---------------------------------------------------------------------------%%
%% joubert.tex
%%---------------------------------------------------------------------------%%

\subsection{Biographical sketch for Wayne Joubert}

\subsubsection*{Address}
\begin{tabular}{l}
  Oak Ridge National Laboratory \\
  Scientific Computing Group\\
  Oak Ridge Leadership Computing Facility\\
  1 Bethel Valley Road, MS-6008\\
  Phone:  (865) 241-2835, Fax:    (865) 281-2850	\\
  Email:  \href{mailto:joubert@ornl.gov}{\texttt{joubert@ornl.gov}}
\end{tabular}

%%---------------------------------------------------------------------------%%

\vspace*{-1ex}
\subsubsection*{Education and Training}
\vspace*{-1ex}
\begin{tabbing}
  \hspace*{1ex} 
  xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx \= xxxxxxxxxxxxxxxxxxxxxxxxx \= xxxxxxxxx \kill
  University of Louisiana \> Mathematics \> B.S.  1981 \\
  Univerity of Texas      \> Mathematics \> Ph.D. 1990
\end{tabbing}

%%---------------------------------------------------------------------------%%

\vspace*{-3ex}
\subsubsection*{Research and Professional Experience}
\vspace*{-1ex}
\begin{tabbing}
  \hspace*{1ex} 
  %%
  \= 2008 -- present \hspace*{2ex} \= Senior R\&D Staff, Oak Ridge National Laboratory\\
  %%
  \> 2007 -- 2008 \> Computational Scientist, Engineer Research and Development Center\\
  %%
  \> 2002 -- 2007 \> Computer Scientist, Xiylon Software\\
  %%
  \> 1999 -- 2001 \> Computer Scientist, LizardTech, Inc.\\
  %%
  \> 1998 -- 1999 \> Team Leader, Los Alamos National Laboratory\\
  %%
  \> 1992 -- 1998 \> Technical Staff, Los Alamos National Laboratory\\
\end{tabbing}

%%---------------------------------------------------------------------------%%

\vspace*{-3ex}
\subsubsection*{Related Publications}
\begin{enumerate}
\parskip = -2pt

\item C. Baker, G. Davidson, T. Evans, S. Hamilton, J. Jarrell and W. Joubert.
   High Performance Radiation Transport Simulations: Preparing for TITAN.
   \textit{International Conference for High Performance Computing, Networking, Storage and Analysis (SC12)}, 2012.

\item W. Joubert and S. Su.
   An Analysis of Computational Workloads on the Jaguar Cray XT System.
   \textit{International Conference on Supercomputing}, 2012.

\item W. Joubert, D. Kothe and H. Nam.
      PREPARING FOR EXASCALE:  ORNL Leadership Computing Application Requirements and Strategy.
   \textit{ORNL/TM-2009/308}, 2009.

\item B. Philip, M. Pernice, W. Joubert and B. Lally.
      A Comparison of Multilevel Preconditioners for Solving Multimaterial Equilibrium Radiation Diffusion Problems on Locally Refined Grids.
   \textit{Nuclear Weapons Highlights}, 2007.

\item W. Joubert and J. Cullum.
   Scalable Algebraic Multigrid on 3500 Processors.
   \textit{Electronic Transactions in Numerical Analysis}, {\bf 23}, 2006.

\item W. Joubert and G. Carey.
   PCG: A Software Package for the Iterative Solution of Linear Systems on Scalar, Vector and Parallel Computers.
   \textit{Scalable High Performance Computing Conference}, 1994.

\item W. Joubert and T. Oppe.
   Improved SSOR and Incomplete Cholesky Solution of Linear Equations on Shared Memory and Distributed Memory Parallel Computers.
   \textit{Numerical Linear Algebra with Applications}, {\bf 1}, 1994.

\item  W. Joubert.
   Lanczos Methods for the Solution of Nonsymmetric Systems of Linear Equations.
   \textit{SIAM Journal on Matrix Analysis and Applications}, {\bf 13}, 1992.
\end{enumerate}

%%---------------------------------------------------------------------------%%

\vspace*{-3ex}
\subsubsection*{Synergistic Activities}
\vspace{-1ex}
%
\begin{enumerate}
\item Performance Task Lead, Oak Ridge Leadership Computing Facility.
\item ORNL Center for Advanced Application Readiness, readiness lead for Denovo code, Titan system.
\item Requirements analysis for 100-250 PF ORNL OLCF-4 system.
\item Member, Society for Industrial and Applied Mathematics.
\end{enumerate}

%%---------------------------------------------------------------------------%%

\vspace*{-3ex}
\subsubsection*{Co-Authors and Collaborators (last 48 months)}
\vspace{-1ex}

{\bf LANL:} David Moulton.
{\bf Mellanox:} Richard Graham.
{\bf Cray:} Jeff Larkin, John Levesque, Norm Troullier.
{\bf ORNL:} Chris Baker, W. Michael Brown, Kevin Clarno, Greg Davidson, Tom Evans, Steven Hamilton, Rebecca Hartman-Baker, Oscar Hernandez, Josh Jarrell, Ricky Kendall, Doug Kothe, Don Maxwell, Bronson Messer, Hai Ah Nam, Bobby Philip, Steve Poole, Ramanan Sankaran, Shiquan Su, Arnold Tharrington, John Turner, Bogdan Vacaliuc.
{\bf SNL:} Richard Barrett.
{\bf University of Colorado:} Samuel Skillman.

\vspace*{-2ex}
\subsubsection*{Graduate Advisor}
\vspace{-1ex}
David M. Young, {\bf University of Texas}.

\pagebreak
\endinput

%%---------------------------------------------------------------------------%%
%% end of joubert.tex
%%---------------------------------------------------------------------------%%

