%%---------------------------------------------------------------------------%%
%% tme.tex
%%---------------------------------------------------------------------------%%

\subsection{Biographical sketch for Thomas M. Evans}

\subsubsection*{Address}
\begin{tabular}{l}
  Oak Ridge National Laboratory \\
  Radiation Transport Group\\
  Reactor and Nuclear Systems Division\\
  One Bethel Valley Road\\
  P.O.~Box 2008, MS-6170\\
  Phone:  (865) 576-3535, Fax:    (865) 574-9619	\\
  Email:  \href{mailto:evanstm@ornl.gov}{\texttt{evanstm@ornl.gov}}
\end{tabular}

%%---------------------------------------------------------------------------%%

\vspace*{-1ex}
\subsubsection*{Education and Training}
\vspace*{-1ex}
\begin{tabbing}
  \hspace*{1ex}
  xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx \= xxxxxxxxxxxxxxxxxxxxxxxxx \= xxxxxxxxx \kill
  Haverford College \> Physics and Astronomy \> B.S. 1992 \\
  Georgia Institute of Technology \> Health Physics \> M.S. 1994 \\
  Georgia Institute of Technology \> Nuclear Engineering \> Ph.D. 1997 
\end{tabbing}

%%---------------------------------------------------------------------------%%

\vspace*{-3ex}
\subsubsection*{Research and Professional Experience}
\vspace*{-1ex}
\begin{tabbing}
  \hspace*{1ex} 
  %%
  \= 2012 -- present \hspace*{2ex} \= Distinguished R\&D Staff, Oak Ridge National
  Laboratory\\ 
  %%
  \> 2007 -- 2012 \> Senior R\&D Staff, Oak Ridge National Laboratory\\
  %%
  \> 2003 -- 2007 \>  Project Leader , Radiation Transport Group,
  Los Alamos National Laboratory\\
  %%
  \> 1997 -- 2003 \> Techical Staff Member, Los Alamos National Laboratory\\
  %%
  \> 1997 -- 1997 \> Post-doctoral Fellow, Los Alamos National Laboratory
\end{tabbing}

%%%---------------------------------------------------------------------------%%

\vspace*{-3ex}
\subsubsection*{Related Publications}
\begin{enumerate}
\parskip = -2pt

\item J.C. Wagner, S.W. Mosher, T.M. Evans, D.E. Peplow, and
  J.A. Turner, {\em Hybrid and Parallel Domain-Decomposition Methods Development to
  Enable Monte Carlo for Reactor Analyses,} {Prog. Nuc. Sci. Tech.},
  {2}:815-820, 2011.
\item G.G. Davidson, T.M. Evans, R.N. Slaybaugh, and C.G. Baker, {\em Massively
  Parallel Solutions to the k-Eigenvalue Problem},
  {Trans. Am. Nucl. Soc.}, {103}, 2010.
\item T.M. Evans, K.T. Clarno, and J.E. Morel, {\em A transport acceleration
  scheme for multigroup discrete ordinates with
  upscattering}, {Nuc. Sci. Eng.}, { 165}:1-13, 2010.
\item T.M. Evans, A.S. Stafford, R.N. Slaybaugh, and K.T. Clarno, 
  {\em Denovo---A
  new three-dimensional parallel discrete ordinates code in
  SCALE}, {Nuc. Tech.}, { 171}:171-200, 2010.
\item T.M. Evans and S.W. Mosher, {\em A Monte Carlo Synthetic
  Acceleration Method for the Non-Linear, Time-Dependent Diffusion
  Equation},  {International Conference on Mathematics,
    Computational Methods and Reactor Physics}, Saratoga Springs, NY,
  ISBN: 978-0-89448-069-0, American Nuclear Society, LaGrange Park,
  IL, 2009.
\item Ryan G. McClarren, Thomas M. Evans, Robert B. Lowrie, and
  Jefferey D. Densmore, {\em Semi-Implicit Time Integration for $P_N$
  Thermal Radiative Transfer},  {J. Comp. Phys.}, 
  { 227}:7561-7586, 2008.
\item J.D. Densmore, T.M. Evans, and M.W. Buksas, {\em A Hybrid
  Transport-Diffusion Algorithm for Monte Carlo Radiation-Transport
  Simulations on Adaptive-Refinement Meshes in $XY$ Geometry},
  {Nuc. Sci. Eng.}, { 159}:1-22, 2008.
\item T.M. Evans and J.D. Densmore {\em Methods for Coupling Radiation,
  Ion, and Electron Energies In Grey Implicit Monte Carlo},
  {J. Comp. Phys.}, { 225}:1695-1720, 2007.
\item T.A. Brunner, T.J. Urbatsch, T.M. Evans, and N.A. Gentile,
  {\em Comparison of Four Parallel Algorithms for Domain Decomposed
  Implicit Monte Carlo}, {J. Comp. Phys.}, { 212}:527--539, 2006.
\item T.M. Evans, T.J. Urbatsch, H.Lichtenstein, and J.E. Morel. A
  residual Monte Carlo Method for Discrete Thermal Radiative
  Diffusion. \textit{J. Comp. Phys.}, {\bf 189}(2), 539--556, 2003.
\end{enumerate}

%%---------------------------------------------------------------------------%%

\vspace*{-3ex}
\subsubsection*{Synergistic Activities}
\vspace{-1ex}
%
\begin{enumerate}
\parskip = -2pt
\item Project Leader, Denovo parallel $S_N$ and {\em Shift} Monte Carlo
  transport codes at Oak Ridge National Laboratory, which are part of the
  SCALE package.
\item PI, Deterministic $S_N$ Transport, CASL Modeling and Simulation Hub at
  Oak Ridge
\item PI, 2012--2013 INCITE Project {\em The Solution of Three-Dimensional PWR
    Neutronics Benchmark Problems for CASL}.
\item Member, technical committee for several Math and Computation Division
  (MCD) meetings of the American Nuclear Society 
\item Officer, MCD Vice President, 2012--2013; incoming MCD President,
  2013--2014; MCD Secretary, 2000--2001.
\end{enumerate}

%%---------------------------------------------------------------------------%%

\vspace*{-3ex}
\subsubsection*{Co-Authors and Collaborators (last 48 months)}
\vspace{-1ex}

{\bf LANL:}  J. Densmore, A.~Hungerford, R.~Lowrie, T.~Urbatsch,
{\bf LLNL:} T.~Brunner, N.~Gentile, 
{\bf Michigan:}  E.~Larsen, B.~Martin, 
{\bf MIT:} B.~Forget,
{\bf ORNL:} C.~Baker, R.~Bartlett, K.~Clarno, G.~Davidson, S.~Hamilton,
J.~Jarrell, S.~Johnson, D.~Kothe, S.~Mosher, D.~Peplow, J.~Turner, J.~Wagner,
C.~Webster, M.~Williams,
{\bf Oregon:} T.~Palmer,
{\bf SNL:} M.~Heroux, R.~Pawlowski
{\bf Texas A\&M:} R.~McClarren, J.~Morel, J.~Ragusa, 
{\bf Wisconsin:}  P.~Wilson

\vspace*{-2ex}
\subsubsection*{Graduate Advisor}
\vspace{-1ex}
C.K.~Wang, {\bf Georgia Institute of Technology}

\vspace*{-2ex}
\subsubsection*{PhD Dissertation Advisees}
\vspace{-1ex}
Rachel Slaybaugh, {\bf University of Wisconsin}
Stuart Slattery, {\bf University of Wisconsin}

\pagebreak
\endinput

%%---------------------------------------------------------------------------%%
%% end of tme.tex
%%---------------------------------------------------------------------------%%

