
% THIS MUST ALL FIT IN 2 PAGES

\subsubsection{Biographical Sketch for Michele Benzi}

% This section is not called out in the RFP, so it could be omitted if
% you are tight on space.
%\vspace*{-1ex}
\subsubsection*{Address}
\vspace*{-1ex}
\begin{tabular}{l}
  Department of Mathematics and Computer Science \\
  Emory University \\
  400 Dowman Drive \\
  Atlanta, GA 30322 \\
  Email: {\tt benzi@mathcs.emory.edu}
\end{tabular}

%\vspace*{-1ex}
\subsubsection*{Education and Training}
\vspace*{-1ex}
\vspace*{-2ex}
\begin{tabbing}
\hspace*{3em}  \= \\
\> Ph.D., 1993, Applied Mathematics, North Carolina State University \\
\> M.S., 1991,  Mathematics, North Carolina State University\\
\> B.S., 1987,  Mathematics, University of Bologna, Italy\\
\end{tabbing}
\vspace*{-4ex}

%\vspace*{-1ex}
\subsubsection*{Research and Professional Experience}
\vspace*{-1ex}
\vspace*{-2ex}
\begin{tabbing}
\hspace*{3em} \= \\
\> Samuel Candler Dobbs Professor (2012--present), Emory University \\ 
\> Professor (2006--present), Emory University\\ 
\> Associate Professor (2000--2006), Emory 
   University \\
\> Staff member (1998--2000), Scientific Computing Group, LANL \\ 
\> Director-funded postdoctoral associate (1997--1998), Scientific 
  Computing Group, LANL\\
\> Postdoctoral Researcher, Parallel Algorithms Group,
  CERFACS, Toulouse, France (1996--1997)\\
\> Researcher, Department of Mathematics, University of Bologna,
   Italy (1993--1996)\\
\end{tabbing}
\vspace*{-4ex}

%\vspace*{-1ex}
\subsubsection*{Publications} 
% List up to ten relevant publications.
%\vspace*{-1ex}
\begin{enumerate*}

\item M.~Benzi and V.~Kuhlemann, ``Restricted additive Schwarz methods
for Markov chains'', Numer.~Linear Algebra Appl., {\bf 18} (2011), 
pp.~1011--1029.

\item M.~Benzi, M.~K.~Ng, W.~Niu and Z.~Wang, ``A relaxed dimensional
factorization preconditioner for the Navier--Stokes equations'', 
J.~Comp.~Phys., {\bf 230} (2011), pp.~6185--6202. 

\item S.~P.~Hamilton, M.~Benzi and E.~Haber, ``New multigrid
smoothers for the Oseen Problem'', Numer.~Linear Algebra Appl., {\bf 17}
(2010), pp.~557--576. 

\item S.~P.~Hamilton, M.~Benzi and J.~S.~Warsa, ``Negative-flux fixups in 
discontinuous finite element S_N transport'', 
Proceedings of the International Conference on Mathematics, 
Computational Methods & Reactor Physics 2009 (M&C 2009), 
American Nuclear Society, Vol. 4 (2009), pp. 2529-2538, 
Saratoga Springs, NY, 2009.

\item M.~Noskov, M.~Benzi, and M.~D.~Smooke, ``An implicit compact 
scheme solver for two-dimensional multicomponent flows'', Computers 
\& Fluids, {\bf 36} (2007), pp.~376--397. 

\item M.~Benzi, G.~H.~Golub, and J.~Liesen, ``Numerical solution of 
saddle point problems'', Acta Numerica, {\bf 14} (2005), pp.~1--137.

\item J.~Warsa, M.~Benzi, T.~Wareing and J.~Morel, ``Preconditioning a 
mixed discontinuous finite element method for radiation diffusion'', 
Numer.~Linear Algebra Appl., {\bf 11} (2004), pp.~795--811. 

\item M.~Benzi, ``Preconditioning techniques for large linear systems:
                  a survey'', J.~Comp.~Phys., {\bf 182} (2002), pp.~418--477.

\item M.~Benzi and M.~Tuma, 
``A parallel solver for large-scale Markov chains'',
                  Appl.~Numer.~Math., {\bf 41} (2002), pp.~135--153. 

\item M.~Benzi, J.~Marin and M.~Tuma, ``A two-level parallel preconditioner
based on sparse approximate inverses'', in Iterative Methods in Scientific 
Computation IV, D. R. Kincaid and A. C. Elster, eds., IMACS Series in 
Computational and Applied Mathematics, Vol.~5, IMACS, New Brunswick, 
NJ (1999), pp.~167--178. 


\end{enumerate*}

%\pagebreak

%\vspace*{-1ex}
\subsubsection*{Synergistic Activities} 
%\vspace*{-1ex}
\begin{enumerate*}
\item Chair, SIAM Activity Group on Linear Algebra (2010-present); 
\item Member
of the SIAM Council (2009-present); 
\item SIAM Fellow (Class of 2012); 
\item Program Committeee Co-chair,
2012 SIAM Applied Linear Conference (Valencia, Spain, June 18-22, 2012); 
\item Program Committeee Co-chair, SIAM Annual Meeting (Minneapolis, MN,
July 9-13, 2012).
\item Editor-in-Charge, SISC Copper Mountain Special Issue (2012-present).
\end{enumerate*}

%\vspace*{-1ex}
\subsubsection*{Collaborators and Co-editors}
% List Collaborators for the past 48 months and co-editors for the
% past 24 months.
\vspace*{-1ex}

{\parindent 0.2in \narrower 
E.~Agichtein (Emory), Z.-Z.~Bai (CAS, Beijing), 
P.~Boito (Limoges), M.~Challacombe (LANL), F.~Chen (Beijing),
E.~Estrada (Strathclyde), L.~Ferragut (Saragoza), 
X.-P.~Guo (Shanghai), E.~Haber (UBC, Canada)
S.~Hamilton (ORNL), N.~Hatano (Tokyo), M.~K.~Ng (Hong Kong), 
Q.~Niu (Zuhai, PRC), M.~Pennacchio (Pavia), M.~Olshanskii (Moscow), 
N.~Razouk (Emory), L.~Rebholz (Clemson), V.~Simoncini 
(Bologna), Y.~Wang (Emory), Z.~Wang (ORNL), Z.-Q.~Wang (Shanghai),
J.~Warsa (LANL).
}

%%\vspace*{1ex}
\subsubsection*{Graduate and Postdoctoral Advisors and Advisees}
% List all advisors and advisees for the past 5 years
\vspace*{-1ex}
%%{\parindent 0.2in \narrower 
Advisors:
\begin{enumerate*}
\item Wayne Joubert (Postdoc, LANL)
\item Iain Duff (Postdoc, CERFACS)
\item Carl Meyer (PhD, NCSU)
\end{enumerate*} 
\vspace*{-1ex}
Advisees: 
\begin{enumerate*}
\item Paola Boito (Postdoc) 
\item Bora Ucar (Postdoc) 
\item Jia Liu (PhD)
\item Nader Razouk (PhD)
\item Lauren Taralli (PhD)
\item Steven Hamilton (PhD)
\item Zhen Wang (PhD)
\item Christine Klymko (PhD)
\item Verena Kuhlemann (PhD)
\item Cheng-yi Zhang (PhD)
\end{enumerate*}
