
The Department of Mathematics and Computer Science at 
Emory University maintains appropriate computing facilities for our
research needs. These include
a shared memory, multi-processor system suited for multi-threaded, 
CPU-intensive programs. The current hardware is a Sun Microsystems 
SunFire V40z, with 4 Dual Core AMD Opteron(tm) Processors and 32 GB of 
memory running Linux.

The department also maintains a second shared memory, multi-processor system 
 suited for multi-threaded, CPU intensive programs. The current hardware 
is a Sun Microsystems SunFire V880, with 8 CPUs and 16 GB of memory running 
Solaris. 

Additionally,
the department maintains {\em Puma},
a high performance cluster with 32 nodes and 
128 processor cores. Each node has two dual core AMD 2214 2.2 GHz 
Opteron CPUs, 4 GB RAM and an 80 GB drive. 
The nodes are connected via a High Performance InfiniBand network and also 
Gigabit Ethernet. The nodes are running Linux CentOS. 

The department also maintains a parallel computing cluster with 
32 CPUs and 16 GB of memory split across 16 identical systems. 
All nodes use RedHat Linux/Intel for their operating system. 
Nodes are interconnected with switched, 100 Mb ethernet. 
Two build environments are available for MPI programs. 
One uses the GNU compiler suite [gcc/g++/g77]; 
the other uses the Portland Group (PGI) compiler suite.

The department also utilizes workstations from the teaching 
lab to form a compute grid for off-hour and weekend use. 
The pool of nodes normally includes 32 identical systems each 
with 2 SPARC CPUs; by arrangement, 16 additional systems could 
be recruited to bring the total CPU count up to 96. 

Emory researchers have also access to a 
256 dual-core, dual-socket AMD Opteron-based compute cluster,
with a total of 1024 CPUs.
Each node uses AMD 2218 processors and has
8 GB DDR2 RAM per node (or 2 GB per core),
250 GB SATA drive local storage per node with 
Gigabit Ethernet interconnect.
The total global storage (parallel file system)
is 8 TB.

Further resources for Emory researchers are available at the
Emerson Center for Scientific Computation. 


